Double White Dwarf Mergers with CASTRO\\ I. Methodology and Code Verification

### 2013-05-07 paper.tex
add acks
add performance outline
more on flattening

The presence of $q_i$ in this expression serves as a reference state.
The idea is that the error in the characteristic projection of the
jump (due to the nonlinearity) is minimized if we pick a suitable
reference state, since we only care about the jumps carried to the
interface over the timestep \citep{colellaglaz1985}.  The original
Castro implementation used the cell-average quantity.  Experiments
found that this is ill-behaved in the presence of strong shocks
(especially at low resolution). For the present work, we switch the
reference state to
\begin{equation}
\tilde{q}_L = \left \{ \begin{array}{cc}
       \mathcal{I}_+^{(+)}(q_i) & \mathrm{if~} u + c > 0 \\
       q_i                    & \mathrm{otherwise}
\end{array}
\right .
\end{equation}
where the $(+)$ superscript here means the fastest wave moving to the right
(the $u+c$ eigenvalue).   This is simply the average under the largest
portion of the parabolic profile that could possible reach the interface 
over the timestep.  This is
in agreement with \citet{ppmunsplit} (eq. 90).  This makes our
expression appear as:
\begin{equation}
q_{i+1/2,L}^{n+1/2} = \tilde{q}_L -
   \sum_{\nu;\lambda_i^{(\nu)}\ge 0} l_i^{(\nu)} \cdot \left [
        \tilde{q_i}  - \mathcal{I}^{(\nu)}_+(q_i)
       \right ] r_i^{(\nu)}
\end{equation}
This change can be used in the public version of Castro by setting
{\tt castro.ppm\_reference = 1}.  Figure~\ref{Fig:sod} shows the
solution to the Sod problem using the old and new reference states.
We see that the shock and contact are slightly sharper with this new
reference state, but also there is a slight dip at the tail of the
rarefaction with the new method.

@@

@@

>>>>

The presence of $q_i$ in this expression serves as a reference state.
The idea is that the error in the characteristic projection of the
jump (due to the nonlinearity) is minimized if we pick a suitable
reference state, since we only care about the jumps carried to the
interface over the timestep \citep{colellaglaz1985}.  The original
Castro implementation used the cell-average quantity.  Experiments
found that this is ill-behaved in the presence of strong shocks
(especially at low resolution). For the present work, we switch the
reference state to
\begin{equation}
\tilde{q}_L = \left \{ \begin{array}{cc}
       \mathcal{I}_+^{(+)}(q_i) & \mathrm{if~} u + c > 0 \\
       q_i                    & \mathrm{otherwise}
\end{array}
\right .
\end{equation}
where the $(+)$ superscript here means the fastest wave moving to the right
(the $u+c$ eigenvalue).   This is simply the average under the largest
portion of the parabolic profile that could possible reach the interface 
over the timestep.  This is
in agreement with \citet{ppmunsplit} (eq. 90).  This makes our
expression appear as:
\begin{equation}
q_{i+1/2,L}^{n+1/2} = \tilde{q}_L -
   \sum_{\nu;\lambda_i^{(\nu)}\ge 0} l_i^{(\nu)} \cdot \left [
        \tilde{q_i}  - \mathcal{I}^{(\nu)}_+(q_i)
       \right ] r_i^{(\nu)}
\end{equation}
The final change is that now we can no longer simply multiply
$\tilde{q_i} - \mathcal{I}^{(\nu)}_+(q_i)$ by the flattening
parameter, $\chi$, but instead blend the traced state with the cell
centered state as:
  \begin{equation}
  q_{i+1/2,\{L,R\}}^{n+1/2} \leftarrow (1 - \chi) q_i + \chi q_{i+1/2,\{L,R\}}^{n+1/2}
  \end{equation}
This is the same prescription used in the \cite{ppm}.

This change can be used in the public version of Castro by setting
{\tt castro.ppm\_reference = 1}.  Figure~\ref{Fig:sod} shows the
solution to the Sod problem using the old and new reference states.
We see that the shock and contact are slightly sharper with this new
reference state, but also there is a slight dip at the tail of the
rarefaction with the new method.

@@

%==========================================================================
% Performance
%==========================================================================
\section{Performance}\label{Sec:Performance}

strong scaling plot

space filling curve?

OMP vs. MPI?

breakdown by component?

@@

This research was supported by NSF award AST-1211563.
This research used resources of the National Energy Research
Scientific Computing Center, which is supported by the Office of
Science of the U.S. Department of Energy under Contract
No. DE-AC02-05CH11231.
